---
title: "The Brain & Mind Centre"
author: "Sydney Informatics Hub"
subtitle: Day 2
output:
  html_document:
    hightlight: github
    theme: lumen
    toc: yes
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = FALSE)
```

## Controls & Loops

These allow you to control the flow of execution of a script typically inside of a function. Common ones include:

* if, else
* for
* while
* repeat
* break
* next
* function

### If loop
```{r, eval=FALSE, message=FALSE}

if (condition) {
    # do something
} else {
    # do something else
}

```

e.g.

```{r}
x=1
if(x > 1) {
  print("x is greater than 1")
  } else {
    print("x is less than 1")
  } 
```

### For loops

A for loop works on an iterable variable and assigns successive values till the end of a sequence.

```{r}
for (i in 1:10) {
    print(i)
}

```

```{r}
x <- c("apples", "oranges", "bananas", "strawberries")

for (i in 1:4) {
    print(x[i])
}

for (i in seq(x)) {
    print(x[i])
}

for (i in 1:4) print(x[i])
```

### While loop
```{r}
i <- 1
while (i < 10) {
    print(i)
    i <- i + 1
}
```

Be sure there is a way to exit out of a while loop.

### Repeat Loops

The repeat loop is an infinite loop and used in association with a break statement.

```{r}
a = 1
repeat { 
  print(a) 
  a = a+1 
  if(a > 4) break }

```

### Break statement

A break statement is used in a loop to stop the iterations and flow the control outside of the loop.

```{r}
x = 1:10 
 for (i in x){ 
     if (i == 2){ 
         break 
     }
     print(i)
 }
```

### Next statement

Next statement enables to skip the current iteration of a loop without terminating it.

```{r}
x = 1: 4 
 for (i in x) { 
     if (i == 2){ 
         next}
     print(i)
 }
```

### Function

If you have to repeat the same few lines of code more than once, then you really need to write a function. Functions are a fundamental building block of R. You use them all the time in R and it's not that much harder to string functions together (or write entirely new ones from scratch) to do more.

* R functions are objects just like anything else.
* By default, R function arguments are lazy - they're only evaluated if they're actually used:
* Every call on a R object is almost always a function call.

#### Basic components of a function

* The body(), the code inside the function.
* The formals(), the "formal" argument list, which controls how you can call the function.
* The `environment()`` which determines how variables referred to inside the function are found.
* args() to list arguments.

```{r}
# Addition 
add <- function(a, b) {
    return(a + b)
}

add(2,5)
```

#### Examples from gambling dataset

```{r,echo=FALSE}
gambling.data <- read.csv(file = "http://data.justice.qld.gov.au/JSD/OLGR/20170817_OLGR_LGA-EGM-data.csv",
                 header = TRUE,
                 sep = ",",
                 stringsAsFactors = FALSE)

names(gambling.data)[2] <- "Local.Govt.Area"

names(gambling.data)[7] <- "Player.Money.Lost"

#Add a day of month (1st) to each date string
date.string <- paste0( "1 " , gambling.data$Month.Year )

#Convert to POSIXlt, a date-time format
strptime( date.string , format = "%d %B %Y" ) -> gambling.data$Date

brisbane.only<-gambling.data[gambling.data$Local.Govt.Area=="BRISBANE",]
row.indicies<-(brisbane.only$Date>="2010-01-01 AEST" &
                 brisbane.only$Date<="2010-12-31 AEST")

(brisbane.2010.data<-brisbane.only[row.indicies,])
```

##### Example 1 : Iterate(Loop) through brisbane.2010.data and sum up the Player.Money.Lost

```{r}
# Iterate(Loop) through brisbane.2010.data and sum up the Player.Money.Lost

brisbane.2010.player.money.lost <- c(brisbane.2010.data$Player.Money.Lost)

total.money.lost = 0

for(i in 1:length(brisbane.2010.player.money.lost)) {
    
  total.money.lost = total.money.lost + brisbane.2010.player.money.lost[i]

}

# Print total amout of money lost
print(total.money.lost)
```


##### Example 2 : Create a function to sum up money lost by passing vector of player.money.lost values

```{r}
# Create a function to sum up money lost by passing vector of player.money.lost values
sum.money.lost <- function(a){
  
  total.money.lost = 0

  # Iterate through vector 
  for(i in 1:length(a)) {
    
    total.money.lost = total.money.lost + a[i]

  }
  return(total.money.lost)
}

# Call function sum.money.lost
sum.money.lost(brisbane.2010.player.money.lost)
```





Adapted from :

https://ramnathv.github.io/pycon2014-r/learn/controls.html

https://www.r-bloggers.com/control-structures-loops-in-r/
