---
title: "The Brain & Mind Centre"
subtitle: "Day 3"
author: "Sydney Informatics Hub"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    theme: lumen
    hightlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = FALSE)
```

## Introduction

Hello this is day 3

### Installation & Setup

blah

```{r installation, eval=FALSE, message=FALSE}
LoadPackages <- function(x){ for( i in x ){ if( ! require( i , character.only = TRUE ) ){
  install.packages( i , dependencies = TRUE , repos = "http://cran.us.r-project.org")
  require( i , character.only = TRUE )}}}

#  Then try/install packages...
LoadPackages(c("tidyverse",
               "xlsx",
               "ggplot2",
               "foreign",
               "dplyr"))

#options(max.print = 100)

options(scipen = 999)
set.seed(5)
```

## About R

### The R Programming Language

[R](https://www.r-project.org/) is an open source programming language and software environment for statistical computing and graphics that is supported by the R Foundation for Statistical Computing. Today, it is one of the most popular languages, being used all across the world in a wide variety of domains and fields.

### RStudio

[RStudio](https://www.rstudio.com/) is an integrated development environment (IDE) for R. It includes a console, syntax-highlighting editor that supports direct code execution, as well as tools for plotting, history, debugging and workspace management.

## Getting Started

### Calculating things in R

Standard math functions work in R:
```{r}
2+3

1e5-5e4

1/1000

sqrt(2)

2*pi
```

Let's use vectors:

```{r}
1:10

2*(1:10)

2^(1:10)
```

We can store values:

```{r}
x<-3
x=3
3->x
x
x^2
```

### Vectors

We can store vectors:

```{r}
x<-1:10
x
2^x

y<-c(3,4,5)
y^2

some.fruits<-c("apple","orange","banana")
some.fruits
```

### Working with Simple Data

A lot of the time in R we are working with tables of data, know as "data frames". 

Commonly, 

**rows** may represent instances e.g. *data points*, *people*, *events*, etc. while 

**columns** will represent different types of data associated with each data point or instance e.g. *Name*, *ID*, *location*, *time*, *value*...

Here is an example data frame

```{r}
simple.data<-data.frame(first.name=c("Alice",
                                     "Bob",
                                     "Cathy",
                                     "Daniel"),
                        gender=as.factor(c("Female",
                                           "Male",
                                           "Female",
                                           "Male")),
                        favourite.number=c(4,23,pi,exp(1)),
                        favourite.letter=c("a","c","x","q"),
                        favourite.weekday=c("Monday","Thursday","Sunday","Friday"),
                        stringsAsFactors = FALSE)
```

